---
title: "p8105_hw2_nh2796"
output: github_document
date: "2024-10-02"
---
##Problem 1

```{r}
library(tidyverse)
subway_data = 
  read_csv(file = "./data/NYC_transit.csv") %>% 
  janitor::clean_names()
print(subway_data)
```
```{r}
subway_data <- subway_data %>% 
  mutate(across(route1:route11, as.character))
```

```{r}
subway_data_tidy = 
  pivot_longer(
    subway_data,
    route1:route11,
    names_to = "routes_served",
    values_to = "train"
  )
```

```{r}
subway_data_tidy <- subway_data_tidy %>% 
  select(-division, -entrance_type, -staffing, -staff_hours, -east_west_street, -ada_notes, -corner, )
```

```{r}
subway_data_tidy <- subway_data_tidy %>% 
  select(-free_crossover, -north_south_street, -exit_only)
```

```{r}
subway_data_tidy <- subway_data_tidy %>% 
  mutate(entry = case_match(
    entry,
    "yes" ~ TRUE,
    "no" ~ FALSE
  ))
```

##
```{r}
distinct_station_count <- subway_data_tidy %>% 
  distinct(station_name) %>% 
  count()
ada_compliant_count <- subway_data_tidy %>% 
  filter(ada == TRUE) %>% 
  distinct(station_name) %>% 
  count()

proportion_entry <- subway_data_tidy %>% 
  filter(vending == "NO") %>% 
  summarise(proportion = mean(entry == TRUE) * 100)
  
 
```

Description of Dataset:
The NYC Transit Dataset provides information regarding the enterances and exits of each subway station. It includes variables that can tell us how many station entrance and exits are ADA compliant, where these stations are located based on latitude and longitude, what routes and trains they serve, and if they have vending or not. In order to clean the data, I first utilized the janitor function to clean the column names and make them more consistent. This data set had many columns that have route numbers. I consolidated these using the pivot longer function. I also removed data columns that were not needed here using the select function. 

Answers to Questions:
There are 356 distinct stations in the NYC transit dataset.
There are 73 distinct stations that are ADA compliant

##problem 2
```{r}
trashwheel_data = 
  readxl::read_excel("./data/Trashwheel.xlsx", sheet = "Mr. Trash Wheel") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(sports_balls)) %>% 
  mutate(sports_balls = as.integer(as.numeric(sports_balls))) %>% 
  select(-x15, -x16) %>% 
  filter(!is.na(dumpster))
 
professor_trashwheel_data =
  readxl::read_excel("./data/Trashwheel.xlsx", sheet = "Professor Trash Wheel") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster))

gwynnda_trashwheel_data = 
  readxl::read_excel("./data/Trashwheel.xlsx", sheet = "Gwynnda Trash Wheel") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster))
```

